---
title: "EDAV Capstone Progress 2"
output: html_document
---

```{r setup, include=FALSE}
library(tidyr)  
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(ggthemes)
library(dplyr)
library(GGally)
library(vcd)
library(extracat)
library(DAAG)
library(forcats)
library(viridisLite)

generic <- read.csv('Data/all_generic_with_districts.csv')

specific <- read.csv('Data/all_specific.csv')

#Create State in specific data
specific <- separate(data = specific, col = district, into = c("state", "district"), sep = "\\-")

specific <- within(specific,  district <- paste(state, district, sep="-"))


genericdf <- as.tibble(generic)
specificdf<-as.tibble(specific)

genericdf <- genericdf %>%
    filter(vote2018!='Will not vote')
specificdf <- specificdf %>%
    filter(vote2018!='Will not vote')
```
Here is a comparison of the datasets across the following demographic variables: Age, Gender, Race, Education, and Marital Status.

```{r}
#############################################################
## Polling v.s. population -- Demographics                 ##
#############################################################
# setwd("D:/Capstone_GitHub/dsi-capstone/")

# create generic demo
age.poll <- generic %>%
  group_by(age) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='generic', Var='Age', Cat=as.numeric(age), Label=age) %>%
  ungroup()
gender.poll <- generic %>%
  group_by(gender) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='generic', Var='Gender', Cat=as.numeric(gender), Label=gender) %>%
  ungroup()
race.poll <- generic %>%
  group_by(race) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='generic', Var='Race', Cat=as.numeric(race), Label=race) %>%
  ungroup()
mar.poll <- generic %>%
  group_by(married) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='generic', Var='Marital Status', Cat=as.numeric(married), Label=married) %>%
  ungroup()
edu.poll <- generic %>%
  group_by(education) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='generic', Var='Education', Cat=as.numeric(education), Label=education) %>%
  ungroup()

generic_demographics <- rbind(age.poll[, 2:7], gender.poll[, 2:7], race.poll[, 2:7],
                           mar.poll[, 2:7], edu.poll[, 2:7])

# create specific demo 
specific <- read.csv("Data/all_specific.csv")

age.spec <- specific %>%
  group_by(age) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='specific', Var='Age', Cat=as.numeric(age), Label=age) %>%
  ungroup()
gender.spec <- specific %>%
  group_by(gender) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='specific', Var='Gender', Cat=as.numeric(gender), Label=gender) %>%
  ungroup()
race.spec <- specific %>%
  group_by(race) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='specific', Var='Race', Cat=as.numeric(race), Label=race) %>%
  ungroup()
mar.spec <- specific %>%
  group_by(married) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='specific', Var='Marital Status', Cat=as.numeric(married), Label=married) %>%
  ungroup()
edu.spec <- specific %>%
  group_by(education) %>%
  summarize(Num=n()) %>%
  mutate(Prop=Num/sum(Num), Type='specific', Var='Education', Cat=as.numeric(education), Label=education) %>%
  ungroup()

specific_demographics <- rbind(age.spec[, 2:7], gender.spec[, 2:7], race.spec[, 2:7],
                              mar.spec[, 2:7], edu.spec[, 2:7])


# create population demo
load("Data/projection_space_national_18-10_02.RData")

age.pop <- pops %>%
  group_by(age) %>%
  summarize(Num=sum(N)) %>%
  mutate(Prop=Num/sum(Num), Type='pop', Var='Age', Cat=as.numeric(age), Label=age) %>%
  ungroup()
gender.pop <- pops %>%
  group_by(gender) %>%
  summarize(Num=sum(N)) %>%
  mutate(Prop=Num/sum(Num), Type='pop', Var='Gender', Cat=as.numeric(gender), Label=gender) %>%
  ungroup()
race.pop <- pops %>%
  group_by(race) %>%
  summarize(Num=sum(N)) %>%
  mutate(Prop=Num/sum(Num), Type='pop', Var='Race', Cat=as.numeric(race), Label=race) %>%
  ungroup()
mar.pop <- pops %>%
  group_by(married) %>%
  summarize(Num=sum(N)) %>%
  mutate(Prop=Num/sum(Num), Type='pop', Var='Marital Status', Cat=as.numeric(married), Label=married) %>%
  ungroup()
edu.pop <- pops %>%
  group_by(education) %>%
  summarize(Num=sum(N)) %>%
  mutate(Prop=Num/sum(Num), Type='pop', Var='Education', Cat=as.numeric(education), Label=education) %>%
  ungroup()
pop_demographics <- rbind(age.pop[, 2:7], gender.pop[, 2:7], race.pop[, 2:7],
                               mar.pop[, 2:7], edu.pop[, 2:7])

# combine all for one plot
plot_demographics <- rbind(pop_demographics, generic_demographics, specific_demographics)

# oops, did something wrong with the category, remap them
plot_demographics <- within(plot_demographics, Cat[Label == '18 - 24'] <- '1')
plot_demographics <- within(plot_demographics, Cat[Label == '25 - 34'] <- '2')
plot_demographics <- within(plot_demographics, Cat[Label == '35 - 44'] <- '3')
plot_demographics <- within(plot_demographics, Cat[Label == '45 - 54'] <- '4')
plot_demographics <- within(plot_demographics, Cat[Label == '> 54'] <- '5')
plot_demographics <- within(plot_demographics, Cat[Label == 'Male'] <- '2')
plot_demographics <- within(plot_demographics, Cat[Label == 'Female'] <- '1')
plot_demographics <- within(plot_demographics, Cat[Label == 'Black'] <- '1')
plot_demographics <- within(plot_demographics, Cat[Label == 'Hispanic'] <- '2')
plot_demographics <- within(plot_demographics, Cat[Label == 'White'] <- '3')
plot_demographics <- within(plot_demographics, Cat[Label == 'Other'] <- '4')
plot_demographics <- within(plot_demographics, Cat[Label == 'Male'] <- '2')
plot_demographics <- within(plot_demographics, Cat[Label == 'Married'] <- '1')
plot_demographics <- within(plot_demographics, Cat[Label == 'Unmarried'] <- '2')
plot_demographics[plot_demographics == 'College'] <- "Bachelors"
plot_demographics[plot_demographics == 'No college'] <- "No Bachelors"
plot_demographics <- within(plot_demographics, Cat[Label == 'Bachelors'] <- '1')
plot_demographics <- within(plot_demographics, Cat[Label == 'No Bachelors'] <- '2')

ggplot(data=plot_demographics, aes(x=as.factor(Label),
    # x=as.factor(paste(Cat, Label, sep='\n')), 
                                   y=Prop,
                                   fill = as.factor(Type),
                                   group=as.factor(Type)))+
                                   # linetype=as.factor(Type))) +
  # geom_point(stat="identity",colour='black') +
    facet_wrap(~Var, scales="free", nrow=1, ncol=6)+
    geom_bar(position='dodge', stat = 'identity')+
    # geom_bar(position = 'dodge', aes(y = Prop, fill= as.factor(Type)))+
  # geom_line(size=0.8) +
  # facet_wrap(~Var, scales="free", nrow=1, ncol=6) +
  theme_bw() +
  # scale_fill_manual(values=c('#1f78b4','#33a02c', '#e31a1c', '#ff7f00', '#8856a7'),guide=FALSE) +
    scale_fill_manual(name = 'Guide1', values=c('#b7b7ff','#A8A8A8','#1fb6b6'))+
    # scale_color_manual(name = 'Guide1', values=c('#19AAD1','#F1D3D3','#008000')) +
    # scale_linetype_manual(name = 'Guide1', values = rep(1:3,3))+
  scale_y_continuous(breaks=c(0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1),
                     labels=c('0%','10%', '20%','30%', '40%', '50%', '60%', '70%', '80%', '90%', '100%')) +
  scale_alpha_manual(values=c(1, .3)) +
  ylab('Proportion') +
  labs(alpha='') +
  theme(legend.position="bottom",
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_text(size=10, angle=45, hjust=1),
        strip.text=element_text(size=10),
        strip.background = element_rect(fill='lightgrey'),
        panel.grid = element_line( size=0.2))

ggsave("plots/demographic_comparison_bar.jpg", width=16, height=8)

```

We'll first analyze the generic, then specific, then compare the two.

```{r}
##### Univariate Distribution Analysis #####
# Gender
g1 <- ggplot(genericdf, aes(x=party)) + geom_bar(aes(fill=party))+
    xlab("Political Party")+
    ylab("Count")+
    ggtitle("Respondent Count by Political Party in Generic Polling")

g1<- g1+scale_fill_manual(values = c("blue", "grey", "red"))
g1
ggsave(filename="Output/CountbyParty_generic.jpg", plot=g1)

g2 <- ggplot(genericdf, aes(x=gender)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Gender")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Gender in Generic Polling")

g2<- g2 +scale_fill_manual(values = c("blue", "grey", "red"))
g2
ggsave(filename="Output/PropbyGender_generic_vote.jpg", plot=g2,  width=16, height=8)

# Race
#Reordered by freq count
genericdf$race <- factor(genericdf$race, levels = c('White', 'Hispanic', 'Other', 'Black'))

g3 <- ggplot(genericdf, aes(x=race)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Race")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Race in Generic Polling")
g3<- g3 +scale_fill_manual(values = c("blue", "grey", "red"))
g3

ggsave(filename="Output/PropbyRace_generic_vote.jpg", plot=g3,  width=16, height=8)

#Education
genericdf$education <- factor(genericdf$education, levels = c('No Bachelors', 'Bachelors'))
g4 <- ggplot(genericdf, aes(x=education)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Education")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Education in Generic Polling")
g4<- g4 +scale_fill_manual(values = c("blue", "grey", "red"))
g4
ggsave(filename="Output/PropbyEdu_generic_vote.jpg", plot=g4, width = 16, height = 8)

#Urbanicity 
g6 <- ggplot(genericdf, aes(x=urbanicity)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Urbanicity")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Race in Generic Polling")
g6

ggsave(filename="Output/Propbyurban_generic_vote.jpg", plot=g6, width = 16, height = 8)

#Marital Status
genericdf$married <- factor(genericdf$married, levels = c('Unmarried', 'Married'))

g7 <- ggplot(genericdf, aes(x=married)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Marital Status")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Marital Status in Generic Polling")
g7<- g7 +scale_fill_manual(values = c("blue", "grey", "red"))
g7
ggsave(filename="Output/CountbyMarried_generic_vote.jpg", plot=g7, width = 16, height = 8)

# Age

g8<- ggplot(genericdf, aes(fct_relevel(age, "> 54", after=4))) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Age")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Age in Generic Polling")
g8<- g8 +scale_fill_manual(values = c("blue", "grey", "red"))
g8

ggsave(filename="Output/Propbyage_generic_vote.jpg", plot=g8, width = 16, height = 8)

#### Specific
g1 <- ggplot(specificdf, aes(x=party)) + geom_bar(aes(fill=party))+
    xlab("Political Party")+
    ylab("Count")+
    ggtitle("Respondent Count by Political Party in Specific Polling")

g1<- g1+scale_fill_manual(values = c("blue", "grey", "red"))
g1
ggsave(filename="Output/CountbyParty_specific.jpg", plot=g1)

# Gender
g2 <- ggplot(specificdf, aes(x=gender)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Gender")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Gender in Specific Polling")
g2<- g2 +scale_fill_manual(values = c("blue", "grey", "red"))
g2
ggsave(filename="Output/PropbyGender_specific_vote.jpg", plot=g2, width = 16, height = 8)

# Race
#Reordered by freq count
specificdf$race <- factor(specificdf$race, levels = c('White', 'Hispanic', 'Other', 'Black'))

g3 <- ggplot(specificdf, aes(x=race)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Race")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Race in Specific Polling")
g3<- g3 +scale_fill_manual(values = c("blue", "grey", "red"))
g3

ggsave(filename="Output/PropbyRace_specific_vote.jpg", plot=g3, width = 16, height = 8)

#Education
specificdf$education <- factor(specificdf$education, levels = c('No Bachelors', 'Bachelors'))

g4 <- ggplot(specificdf, aes(x=education)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Education")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Education in Specific Polling")
g4<- g4 +scale_fill_manual(values = c("blue", "grey", "red"))

ggsave(filename="Output/PropbyEdu_specific_vote.jpg", plot=g4, width = 16, height = 8)

#Urbanicity 
g6 <- ggplot(specificdf, aes(x=urbanicity)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Gender")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Urbanicity in Specific Polling")
g6<- g6 +scale_fill_manual(values = c("blue", "grey", "red"))

ggsave(filename="Output/Propbyurban_specific_vote.jpg", plot=g6, width = 16, height = 8)

#Marital Status
specificdf$married <- factor(specificdf$married, levels = c('Unmarried', 'Married'))
g7 <- ggplot(specificdf, aes(x=married)) +
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Marital Status")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Marital Status in Specific Polling")
g7<- g7 +scale_fill_manual(values = c("blue", "grey", "red"))

ggsave(filename="Output/PropbyMarried_specific_vote.jpg", plot=g7, width = 16, height = 8)

# Age
g8 <- ggplot(specificdf, aes(fct_relevel(age, "> 54", after=4)))+
    geom_bar(position='dodge', aes(y = (..count..)/sum(..count..), fill=vote2018))+
    xlab("Age")+
    ylab("Proportion")+
    ggtitle("Proportion of Responses by Age in Specific Polling")
g8<- g8 +scale_fill_manual(values = c("blue", "grey", "red"))

ggsave(filename="Output/Propbyage_specific_vote.jpg", plot=g8, width = 16, height = 8)
```
We'll now translate some of the old Python code to visualize proportions of voting for the majority party across the United States.

```{r}
#Insert translated Geomap here.
library(plotly)
library(RColorBrewer)

## Create version of genericdf and specificdf that have the following columns
## state, total number of responses, prop_rep, prop_dem, prop_other

#BASELINE df 
genericdf <- genericdf %>%
    filter(vote2018!='Other candidate/not sure')
specificdf <- specificdf %>%
    filter(vote2018!='Other candidate/not sure')

generic_map_df <- genericdf %>%
    select(state, vote2018)%>%
  group_by(state)%>%
    mutate(total_num_responses = n())%>%
    filter(vote2018 == 'Republican candidate')%>%
    mutate(num_rep = n())%>%
    select(state, total_num_responses, num_rep) %>%
    distinct(state, total_num_responses, num_rep) %>%
    arrange(state)

dem <- genericdf %>%
    group_by(state)%>%
    filter(vote2018=='Democratic candidate') %>%
    mutate(num_dem = n())%>%
    select(state, num_dem) %>%
    distinct(state, num_dem) %>% 
    arrange(state)

# other <- genericdf %>%
#     group_by(state) %>%
#     filter(vote2018=='Other candidate/not sure') %>%
#     mutate(num_other = n())%>%
#     select(state, num_other) %>%
#     distinct(state, num_other)%>%
#     arrange(state)

generic_map_df <- cbind(generic_map_df, dem)%>%
    select(state, total_num_responses, num_rep, num_dem) %>%
    group_by(state) %>%
    mutate(prop_rep = num_rep / total_num_responses) %>% 
    mutate(prop_dem = num_dem / total_num_responses) 

#BASELINE specificdf 
# Note that the total number of states gathered in 40 districts surveys = 18 states
specific_map_df <- specificdf %>%
    select(state, vote2018)%>%
  group_by(state)%>%
    mutate(total_num_responses = n())%>%
    filter(vote2018 == 'Republican candidate')%>%
    mutate(num_rep = n())%>%
    select(state, total_num_responses, num_rep) %>%
    distinct(state, total_num_responses, num_rep) %>%
    arrange(state)

dem <- specificdf %>%
    group_by(state)%>%
    filter(vote2018=='Democratic candidate') %>%
    mutate(num_dem = n())%>%
    select(state, num_dem) %>%
    distinct(state, num_dem) %>% 
    arrange(state)

# other <- specific %>%
#     group_by(state) %>%
#     filter(vote2018=='Other candidate/not sure') %>%
#     mutate(num_other = n())%>%
#     select(state, num_other) %>%
#     distinct(state, num_other)%>%
#     arrange(state)

specific_map_df <- cbind(specific_map_df, dem)%>%
    select(state, total_num_responses, num_rep, num_dem) %>%
    group_by(state) %>%
    mutate(prop_rep = num_rep / total_num_responses) %>% 
    mutate(prop_dem = num_dem / total_num_responses)

########### Design Choropleth map ###################

generic_map_df$hover <- with(generic_map_df, paste(state,'<br>',
                                         'Number of Responses',total_num_responses,'<br>',
                                         'Proportion likely to vote Republican',prop_rep, '<br>',
                                         'Proportion likely to vote Democrat', prop_dem,  '<br>'))
                                         # 'Proportion likely to vote Other/Not sure',prop_other,'<br>'))

# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
geo <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

colors = rev(brewer.pal(4, "RdYlBu"))
p <- plot_geo(generic_map_df, locationmode = 'USA-states') %>%
  add_trace(
    z = ~prop_rep, text = ~hover, locations = ~state,
    color = ~prop_rep, colors = colors  ) %>%
  colorbar(title = "% of Respondents likely to vote Republican", limits=c(0,1)) %>%
  layout(
    title = 'Which party would you vote for? (Generic Poll)',
    geo = geo
  )


###################### Design choropleth for specific #################
########### Design Choropleth map ###################

specific_map_df$hover <- with(specific_map_df, paste(state,'<br>',
                                         'Number of Responses',total_num_responses,'<br>',
                                         'Proportion likely to vote Republican',prop_rep, '<br>',
                                         'Proportion likely to vote Democrat', prop_dem,  '<br>'))
                                         # 'Proportion likely to vote Other/Not sure',prop_other,'<br>'))

# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
geo <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

colors = rev(brewer.pal(4, "RdYlBu"))
p2 <- plot_geo(specific_map_df, locationmode = 'USA-states') %>%
  add_trace(
    z = ~prop_rep, text = ~hover, locations = ~state,
    color = ~prop_rep, colors = colors  ) %>%
  colorbar(title = "% of Respondents likely to vote Republican", limits=c(0,1)) %>%
  layout(
    title = 'Whom (candidate names mentioned) who would you vote for? (Specific Poll)',
    geo = geo
  )
p2
```
Now we'll transition to making district level plots
