---
title: "stanMRP"
author: "Tin Oreskovic"
date: "11/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nnet)
library(rstanarm)
library(lme4)
library(optimx)
```

## ANEW
```{r generic}
set.seed(42)

demographics <- c('party', 'gender', 'age', 'race', 'education', 'urbanicity', 'married')
demographic_combinations <- c(combn(demographics, simplify=FALSE, 1),
                              combn(demographics, simplify=FALSE, 2),
                              combn(demographics, simplify=FALSE, 3),
                              combn(demographics, simplify=FALSE, 4),
                              combn(demographics, simplify=FALSE, 5),
                              combn(demographics, simplify=FALSE, 6),
                              combn(demographics, simplify=FALSE, 7),
                              'district')

pops <- read_csv('Data/population_data.csv',
                 col_types=cols(age = col_factor(levels=c('18 - 24', '25 - 34', '35 - 44', '45 - 54', '> 54')),
                                urbanicity = col_factor(levels=c('R1', 'R2', 'S3', 'S4', 'U5', 'U6')),
                                gender = col_factor(levels=c('Male', 'Female')),
                                district = col_factor(NULL),
                                state = col_skip(),
                                race = col_factor(levels=c('White', 'Hispanic', 'Black', 'Other')),
                                education = col_factor(levels=c('No Bachelors', 'Bachelors')),
                                married = col_factor(levels=c('Unmarried', 'Married')),
                                party = col_factor(levels=c('Dem', 'Rep', 'Ind')),
                                N = col_double()))

generic_s <- read_csv('Data/all_generic_with_districts.csv',
                    col_types=cols(party = col_factor(levels=levels(pops$party)),
                                   gender = col_factor(levels=levels(pops$gender)),
                                   age = col_factor(levels=levels(pops$age)),
                                   race = col_factor(levels=levels(pops$race)),
                                   education = col_factor(levels=levels(pops$education)),
                                   urbanicity = col_factor(levels=levels(pops$urbanicity)),
                                   vote2018 = col_character(),
                                   married = col_factor(levels=levels(pops$married)),
                                   district = col_factor(levels=levels(pops$district)),
                                   state = col_skip(),
                                   wave = col_skip())) %>%
  mutate(vote2018 = case_when(vote2018 == 'Democratic candidate' ~ 'Democrat',
                              vote2018 == 'Republican candidate' ~ 'Republican')) %>%
  filter(!is.na(vote2018) & !is.na(district)) %>%
  mutate(vote2018 = factor(vote2018, levels=c('Democrat', 'Republican'))) %>%
  sample_n(6209)

pred_pops <- pops %>% group_by_if(is.factor) %>% summarise(N=sum(N))

#generic_m <- glmer(vote2018 ~ age + urbanicity + gender + race + education + married + party + (1 | district), data = generic_s, family = binomial(link = "logit"), control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))

generic_m <- glmer(vote2018 ~ age + urbanicity + gender + race + education + married + party + (1 | district),
                   data = generic_s, family = binomial(link = "logit"),
                   control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))

pred <- predict(generic_m, newdata = pred_pops, type="response", allow.new.levels = TRUE)

pred_pops$pred_val = pred_pops$N * pred

df <- list(Demographic_Type = 'National', Demographic = 'National',
           Vote_R = sum(pred_pops$pred_val)/sum(pred_pops$N)) %>% as.tibble()
for(i in 1:length(demographic_combinations)){
  vars = demographic_combinations[[i]]
  temp = pred_pops %>% group_by_(.dots = lapply(vars, as.symbol)) %>%
    summarise(Vote_R = sum(pred_val) / sum(N)) %>%
    unite_(col='Demographic', vars, sep=' ') %>%
    mutate(Demographic_Type = str_to_title(str_flatten(vars, collapse=' ')))
  
  df <- bind_rows(df, temp)
}

write_csv(df, 'Output/generic_mrp_results_district.csv')
```


## VARYING SLOPES GENERIC 
"Parameters or bounds appear to have different scalings.
  This can cause poor performance in optimization. 
  It is important for derivative free methods like BOBYQA, UOBYQA, NEWUOA.unable to evaluate scaled gradientModel failed to converge: degenerate  Hessian with 1 negative eigenvalues"
```{r generic_var_slopes}
set.seed(42)

demographics <- c('party', 'gender', 'age', 'race', 'education', 'urbanicity', 'married')
demographic_combinations <- c(combn(demographics, simplify=FALSE, 1),
                              combn(demographics, simplify=FALSE, 2),
                              combn(demographics, simplify=FALSE, 3),
                              combn(demographics, simplify=FALSE, 4),
                              combn(demographics, simplify=FALSE, 5),
                              combn(demographics, simplify=FALSE, 6),
                              combn(demographics, simplify=FALSE, 7),
                              'district')

pops <- read_csv('Data/population_data.csv',
                 col_types=cols(age = col_factor(levels=c('18 - 24', '25 - 34', '35 - 44', '45 - 54', '> 54')),
                                urbanicity = col_factor(levels=c('R1', 'R2', 'S3', 'S4', 'U5', 'U6')),
                                gender = col_factor(levels=c('Male', 'Female')),
                                district = col_factor(NULL),
                                state = col_skip(),
                                race = col_factor(levels=c('White', 'Hispanic', 'Black', 'Other')),
                                education = col_factor(levels=c('No Bachelors', 'Bachelors')),
                                married = col_factor(levels=c('Unmarried', 'Married')),
                                party = col_factor(levels=c('Dem', 'Rep', 'Ind')),
                                N = col_double()))

generic_s <- read_csv('Data/all_generic_with_districts.csv',
                    col_types=cols(party = col_factor(levels=levels(pops$party)),
                                   gender = col_factor(levels=levels(pops$gender)),
                                   age = col_factor(levels=levels(pops$age)),
                                   race = col_factor(levels=levels(pops$race)),
                                   education = col_factor(levels=levels(pops$education)),
                                   urbanicity = col_factor(levels=levels(pops$urbanicity)),
                                   vote2018 = col_character(),
                                   married = col_factor(levels=levels(pops$married)),
                                   district = col_factor(levels=levels(pops$district)),
                                   state = col_skip(),
                                   wave = col_skip())) %>%
  mutate(vote2018 = case_when(vote2018 == 'Democratic candidate' ~ 'Democrat',
                              vote2018 == 'Republican candidate' ~ 'Republican')) %>%
  filter(!is.na(vote2018)) %>%
  mutate(vote2018 = factor(vote2018, levels=c('Democrat', 'Republican'))) %>%
  sample_n(6209)

pred_pops <- pops %>% group_by_if(is.factor) %>% summarise(N=sum(N))

generic_m2 <- glmer(vote2018 ~ age + urbanicity + gender + race + education + married + party + age*gender + (1 + age + gender | district),
                    data = generic_s, family = binomial(link = "logit"), control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))

#generic_m2 <- glmer(vote2018 ~ age + urbanicity + gender + race + education + married + party + (1 + age + gender | district), data = generic_s, family = binomial(link = "logit"))

pred <- predict(generic_m2, newdata = pred_pops, type= "response", allow.new.levels = TRUE)

pred_pops$pred_val = pred_pops$N * pred

df <- list(Demographic_Type = 'National', Demographic = 'National',
           Vote_R = sum(pred_pops$pred_val)/sum(pred_pops$N)) %>% as.tibble()
for(i in 1:length(demographic_combinations)){
  vars = demographic_combinations[[i]]
  temp = pred_pops %>% group_by_(.dots = lapply(vars, as.symbol)) %>%
    summarise(Vote_R = sum(pred_val) / sum(N)) %>%
    unite_(col='Demographic', vars, sep=' ') %>%
    mutate(Demographic_Type = str_to_title(str_flatten(vars, collapse=' ')))
  
  df <- bind_rows(df, temp)
}

write_csv(df, 'Output/generic_mrp_results_district_slopes.csv')
```


# SPECIFIC

```{r specific}
pops <- read_csv('Data/population_data.csv',
                 col_types=cols(age = col_factor(levels=c('18 - 24', '25 - 34', '35 - 44', '45 - 54', '> 54')),
                                urbanicity = col_factor(levels=c('R1', 'R2', 'S3', 'S4', 'U5', 'U6')),
                                gender = col_factor(levels=c('Male', 'Female')),
                                district = col_factor(NULL),
                                state = col_skip(),
                                race = col_factor(levels=c('White', 'Hispanic', 'Black', 'Other')),
                                education = col_factor(levels=c('No Bachelors', 'Bachelors')),
                                married = col_factor(levels=c('Unmarried', 'Married')),
                                party = col_factor(levels=c('Dem', 'Rep', 'Ind')),
                                N = col_double()))

specific_s <- read_csv('Data/all_specific.csv',
                     col_types=cols(party = col_factor(levels=levels(pops$party)),
                                    gender = col_factor(levels=levels(pops$gender)),
                                    age = col_factor(levels=levels(pops$age)),
                                    race = col_factor(levels=levels(pops$race)),
                                    education = col_factor(levels=levels(pops$education)),
                                    urbanicity = col_factor(levels=levels(pops$urbanicity)),
                                    married = col_factor(levels=levels(pops$married)),
                                    vote2018 = col_character(),
                                    district = col_factor(levels = levels(pops$district)),
                                    wave = col_skip())) %>%
  mutate(vote2018 = case_when(vote2018 == 'Democratic candidate' ~ 'Democrat',
                              vote2018 == 'Republican candidate' ~ 'Republican')) %>%
  filter(!is.na(vote2018)) %>%
  mutate(vote2018 = factor(vote2018, levels=c('Democrat', 'Republican')))


pred_pops <- pops %>% group_by_if(is.factor) %>% summarise(N=sum(N))


specific_m <- glmer(vote2018 ~ age + urbanicity + gender + race + education + married + party + (1 | district),
                    data = specific_s, family = binomial(link = "logit"),
                    control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))

pred <- predict(specific_m, newdata=pred_pops, type="response", allow.new.levels = TRUE)

pred_pops$pred_val = pred_pops$N * pred

df <- list(Demographic_Type = 'National', Demographic = 'National',
           Vote_R = sum(pred_pops$pred_val)/sum(pred_pops$N)) %>% as.tibble()
for(i in 1:length(demographic_combinations)){
  vars = demographic_combinations[[i]]
  temp = pred_pops %>% group_by_(.dots = lapply(vars, as.symbol)) %>%
    summarise(Vote_R = sum(pred_val) / sum(N)) %>%
    unite_(col='Demographic', vars, sep=' ') %>%
    mutate(Demographic_Type = str_to_title(str_flatten(vars, collapse=' ')))
  
  df <- bind_rows(df, temp)
}

write_csv(df, 'Output/specific_mrp_results_district.csv')
```


# VARYING SLOPES SPECIFIC
"Parameters or bounds appear to have different scalings.
  This can cause poor performance in optimization. 
  It is important for derivative free methods like BOBYQA, UOBYQA, NEWUOA."
```{r specific_var_slopes}
pops <- read_csv('Data/population_data.csv',
                 col_types=cols(age = col_factor(levels=c('18 - 24', '25 - 34', '35 - 44', '45 - 54', '> 54')),
                                urbanicity = col_factor(levels=c('R1', 'R2', 'S3', 'S4', 'U5', 'U6')),
                                gender = col_factor(levels=c('Male', 'Female')),
                                district = col_factor(NULL),
                                state = col_skip(),
                                race = col_factor(levels=c('White', 'Hispanic', 'Black', 'Other')),
                                education = col_factor(levels=c('No Bachelors', 'Bachelors')),
                                married = col_factor(levels=c('Unmarried', 'Married')),
                                party = col_factor(levels=c('Dem', 'Rep', 'Ind')),
                                N = col_double()))

specific_s <- read_csv('Data/all_specific.csv',
                     col_types=cols(party = col_factor(levels=levels(pops$party)),
                                    gender = col_factor(levels=levels(pops$gender)),
                                    age = col_factor(levels=levels(pops$age)),
                                    race = col_factor(levels=levels(pops$race)),
                                    education = col_factor(levels=levels(pops$education)),
                                    urbanicity = col_factor(levels=levels(pops$urbanicity)),
                                    married = col_factor(levels=levels(pops$married)),
                                    vote2018 = col_character(),
                                    district = col_factor(levels = levels(pops$district)),
                                    wave = col_skip())) %>%
  mutate(vote2018 = case_when(vote2018 == 'Democratic candidate' ~ 'Democrat',
                              vote2018 == 'Republican candidate' ~ 'Republican')) %>%
  filter(!is.na(vote2018)) %>%
  mutate(vote2018 = factor(vote2018, levels=c('Democrat', 'Republican')))

pred_pops <- pops %>% group_by_if(is.factor) %>% summarise(N=sum(N))


specific_m2 <- glmer(vote2018 ~ age + urbanicity + gender + race + education + married + party + (1 + age + gender | district),
                     data = specific_s, family = binomial(link = "logit"),
                     control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))

pred <- predict(specific_m2, newdata=pred_pops, type="response", allow.new.levels = TRUE)

pred_pops$pred_val = pred_pops$N * pred

df <- list(Demographic_Type = 'National', Demographic = 'National',
           Vote_R = sum(pred_pops$pred_val)/sum(pred_pops$N)) %>% as.tibble()
for(i in 1:length(demographic_combinations)){
  vars = demographic_combinations[[i]]
  temp = pred_pops %>% group_by_(.dots = lapply(vars, as.symbol)) %>%
    summarise(Vote_R = sum(pred_val) / sum(N)) %>%
    unite_(col='Demographic', vars, sep=' ') %>%
    mutate(Demographic_Type = str_to_title(str_flatten(vars, collapse=' ')))
  
  df <- bind_rows(df, temp)
}

write_csv(df, 'Output/specific_mrp_results_district_slopes.csv')
```








## IF WE EVER WANT THE NESTED MODELS:

```{r nested}
pops2 <- read_csv('Data/population_data.csv',
                 col_types=cols(age = col_factor(levels=c('18 - 24', '25 - 34', '35 - 44', '45 - 54', '> 54')),
                                urbanicity = col_factor(levels=c('R1', 'R2', 'S3', 'S4', 'U5', 'U6')),
                                gender = col_factor(levels=c('Male', 'Female')),
                                district = col_factor(NULL),
                                state = col_skip(),
                                race = col_factor(levels=c('White', 'Hispanic', 'Black', 'Other')),
                                education = col_factor(levels=c('No Bachelors', 'Bachelors')),
                                married = col_factor(levels=c('Unmarried', 'Married')),
                                party = col_factor(levels=c('Dem', 'Rep', 'Ind')),
                                N = col_double()))

specific_s <- read_csv('Data/all_specific.csv',
                     col_types=cols(party = col_factor(levels=levels(pops2$party)),
                                    gender = col_factor(levels=levels(pops2$gender)),
                                    age = col_factor(levels=levels(pops2$age)),
                                    race = col_factor(levels=levels(pops2$race)),
                                    education = col_factor(levels=levels(pops2$education)),
                                    urbanicity = col_factor(levels=levels(pops2$urbanicity)),
                                    married = col_factor(levels=levels(pops2$married)),
                                    vote2018 = col_character(),
                                    district = col_factor(levels = levels(pops2$district)),
                                    wave = col_skip())) %>%
  mutate(vote2018 = case_when(vote2018 == 'Democratic candidate' ~ 'Democrat',
                              vote2018 == 'Republican candidate' ~ 'Republican')) %>%
  filter(!is.na(vote2018)) %>%
  mutate(vote2018 = factor(vote2018, levels=c('Democrat', 'Republican')),
         state = str_sub(district, 0, 2))

pred_pops2 <- pops2 %>% group_by_if(is.factor) %>% summarise(N=sum(N))


specific_m2 <- glmer(vote2018 ~ age + urbanicity + gender + race + education + married + party + (1 | district),
                     data = specific_s, family = binomial(link = "logit"),
                     control = glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')))
pred2 <- predict(specific_m2, newdata=pred_pops2, type="response", allow.new.levels = TRUE)


pred_pops2$pred_val = pred_pops2$N * pred2

df2 <- list(Demographic_Type = 'National', Demographic = 'National',
           Vote_R = sum(pred_pops2$pred_val)/sum(pred_pops2$N)) %>% as.tibble()
for(i in 1:length(demographic_combinations)){
  vars = demographic_combinations[[i]]
  temp = pred_pops2 %>% group_by_(.dots = lapply(vars, as.symbol)) %>%
    summarise(Vote_R = sum(pred_val) / sum(N)) %>%
    unite_(col='Demographic', vars, sep=' ') %>%
    mutate(Demographic_Type = str_to_title(str_flatten(vars, collapse=' ')))
  
  df2 <- bind_rows(df2, temp)
}

write_csv(df2, 'Data/specific_mrp_results_nested2.csv')
```
